---
title: "ALB analysis"
format:
  html:
    self-contained: true
    code-fold: false
    code-summary: "Code details"
    toc: true
    toc-location: left
    toc-depth: 3
    toc-expand: 2
  pdf:
    pdf-engine: lualatex
    toc: false
    number-sections: true
---

## The ALB data

The ALB are in the latest version  `6.0-10` of the `gamlss.data` package. (Note that we may have to get the latest version from the [gamlss universe](https://gamlss-dev.r-universe.dev/builds). 



Getting the data from `gamlss.data`;

```{r}
#| warning: false
library(gamlss)
data(ALB)
dim(ALB)
head(ALB)
```

Now we plot the data;

```{r}
plot(ALB~ga, data=ALB)
```

The data were collected when  healthy pregnant woman had their scans at Kings College hospital. The scan are taken routinely four times. That explains why they are caps in data for different  gestational age. Notice that ate this data they are NOT repeated observations, all 3429 women are different.

::: callout-tip
## Purpose

The purpose of the study is to try to identify patients with low `ALB`
:::

## Fit a model

Here we are using a `distributional regression` model, GAMLSS, to answer the question.In a distributional model all the parameters of the distributions are functions of the explanatory variable gestation age, `ga`.  

Any GAMLSS regression model needs two type of assumtions to be fitted:

i) the assumed `distribution` for the response i.e `family=GA`,

ii) the `selection` of covariates;  In fact in general since we are modelling all distribution parameters as functions of the covariates, we are looking `which covariate` affect `which distribution parameter`.


To start a selection procedure in GAMLSS we need a starting model.

Here we fit  two starting models; 

i) a normal model e.g. `family=NO` which works as a staring model for the function  `chooseDist()`


```{r}
#| cache: true
m00 <- gamlss(   ALB ~ pb(ga),
    sigma.formula = ~ pb(ga),
      nu.formula = ~ pb(ga),  
     tau.formula = ~ pb(ga),
             family = NO, data = ALB) 
```

ii) a BCTo model e.g. `family=BCTo` ` which  is used as staring model for the function  `chooseDistCV()` (see note below for an explanation why we have to do that). The function  `chooseDistCV()`is a new function and has to be sourced.


```{r}
#| cache: false
m0 <- gamlss(   ALB ~ pb(ga),
    sigma.formula = ~ pb(ga),
      nu.formula = ~ pb(ga),  
     tau.formula = ~ pb(ga),
             family = BCTo, data = ALB, n.cyc=100) 
```

::: callout-note
The `nu.formula = ~ pb(ga)` and `tau.formula = ~ pb(ga)` will be not be used by  `chooseDistCV()` unless more that two parameter are involved in the distribution. This has to be investigated further. 
:::


::: callout-tip
## reminder

I have to find out why  the function `chooseDistCV()` do not accepts the normal error model like  `chooseDist()`. 
:::


## Finding a Distributions (using the training data)  

The typical function for finding a distribution is `chooseDist()`. 
The problem is that the function only finds suitable distributions for the training data  using  an Generalized AKaike information criterion,  GAIC,  as the `training  measure of goodness of fit`. Therefor it do not check for over-fitting.

```{r}
#| cache: true
D1=chooseDist(m00, data=ALB, type="realplus", parallel="snow", ncpus = 10)
```
With penalties $k=2$ and $k=2.84$ the `BCT` is coming best while for BIC the `BCPEo`.  

Next we print a table of all the fitted GAIC;
```{r}
library(knitr)
kable(D1)
```
The problem with the above analysis is that the measure of goodness of fit, `GAIC`, was validated on the training data only This show that the model is probably adequate but it do not check for overfitting.


## Finding a Distributions (cross validation)  

Next we use a $10^{th}$ fold  cross validation for choosing the distribution. Cross validation hopefully prevent overfiting.

There are two criteria we used here;  

- the  `test deviance` which is a `global` criterion, that is, it evaluates how the distribution fits overall.

- a `chi-square` criterion which `focus` on the lower centiles of the distribution.


The new function `chooseDistCV()` is used which is not in the `gamlss` package (yet) so we have to source it;

```{r}
source("chooseDistCVParallel.R")
```

### test deviance


First we use a  global measurement of goodness of fit criterion the `test deviance`. That is -2 the log-Likelihood  evaluated at the test data rather that the training ones. I order to do that we have to define the function  `measure_of_goodness_of_fit()` which  `chooseDistcCV()` uses  to evaluate the  different distributions. Here we define the function;


```{r}
measure_of_goodness_of_fit <- function(object)
{
  object$CV
}
```

Now to avoid fitting all possible positive range distributions using cross validation we speed the process  by using only  a selected set of 5 distributions. The gamma., `GA` and the log-normal, `LOGNO`, distributions both very popular in the statistical literature and the distributions `BCTo`, `BCPEo` and `BCCGo` popular in centile estimation (associated with the LMS method).     The `R` code is as follows; 

```{r}
#| cache: true

PP = chooseDistCV(m0, data=ALB, type="extra", 
          extra=c("GA", "LOGNO", "BCCGo", "BCPEo", "BCTo"))
```
we can print the order of the fit using;
```{r}
PP
```


The simple use of the  `chooseDistcCV()` function results to output which prints the fold number after a new model is fitted to each partitio. One could avoid the unnecessary output by using parallel processing.   

```{r}
#| cache: true
set.seed(123)
PP1 = chooseDistCV(m0, data=ALB, type="extra", 
          extra=c("GA", "LOGNO", "BCCGo", "BCPEo", "BCTo"),
          parallel="snow", ncpus = 10)
```

```{r}
PP1
```


::: callout-note

There are few things to note  here; 

- Any cross validation  will have results depending on the original computer seed 

- Using parallel computation the original seed do not apply in all CPUS therefore the results are in general different.  
:::

Fortunately the `BCTo` distribution gives the best fit in both runs.  This is not always the case.

::: callout-tip
## checking the code 

The code below is used to find out whether using the factor which creates the partition using the argument `rand` to generate the partitions makes any diffence.  It looks that it does not.  
:::


```{r}
#| cache: true
set.seed(123)
sam <- sample(10 , dim(ALB)[1], replace=TRUE)
PP <- chooseDistCV(m0, data=ALB, type="extra",
          extra=c("GA", "LOGNO", "BCCGo", "BCPEo", "BCTo"), rand=sam)
PP1 <- chooseDistCV(m0, data=ALB, type="extra", 
           extra=c("GA", "LOGNO", "BCCGo", "BCPEo", "BCTo"),
           parallel="snow", ncpus = 10, rand=sam)
```

```{r}
 PP 
 PP1
```

::: callout-tip
## Finish checking the code 

It looks that the the option `rand=sam` is NOT working properly in the `chooseDistCV()`
  
:::


### chi-squared (the focus measure )


First bring the functions in;
```{r}
source("Chi2.R")
```

Define the focus measure of goodness of fit; 
```{r}
measure_of_goodness_of_fit <- function(object)
 {
   cal_chi2(object$residCV, xvar=ALB$ga, breaks = c(0,17,25,34.5, 50), tail.percent = 5,  side= "lower")
 }
```
Now choose distribution using CV with no parallel computaions.
```{r}
#| cache: true
 PPP = chooseDistCV(m0, data=ALB, type="extra",
           extra=c("GA", "LOGNO", "BCCGo", "BCPEo", "BCTo"))
```

Now choose distribution using CV with parallel computaions.
```{r}
#| cache: true
PPP1 = chooseDistCV(m0, data=ALB, type="extra", 
          extra=c("GA", "LOGNO", "BCCGo", "BCPEo", "BCTo"),
          parallel="snow", ncpus = 10)
```
Now print the results
```{r}
PPP
PPP1
```

::: callout-tip


## reminder

This interesting the same sellection procedure points to a different distribution
:::
